---
title: "Week9 Assignment"
author: "Senthil Dhanapal"
date: "Saturday, March 28, 2015"
output: html_document
---

### I tried to access raw version of my files from github to load it directly from a remote server but my computer had issues with some certificates. I tried to address it but I couldn't resolve it. So, I put the code as comment.

### load required packages
```{r, echo=FALSE, warning=FALSE}
#install.packages("XML")
#install.packages("RCurl")
#install.packages("plyr")
#install.packages("jsonlite")
library(XML)
library(plyr)
#library(RCurl)
library(jsonlite)
library(dplyr)


```


### books.html with a "field=value" format
```{r}

books.html.url <- "C:/Senthil/MSDataAnalytics/Semester1/Assignments/IS607/Week9Assignment/books.html"
#books.html.url <- getURL("https://raw.githubusercontent.com/senthiltamil/CUNY_MSDA/master/Week9_Assignment/books.html")
books.html.data <- readHTMLTable(books.html.url,header = F)
books.html.data.df <- as.data.frame(books.html.data)
str(books.html.data.df)
print(books.html.data.df)

```


### books.html with a table-row format
```{r}

#books.html.url <- getURL("https://raw.githubusercontent.com/senthiltamil/CUNY_MSDA/master/Week9_Assignment/books1.html")
books1.html.url <- "C:/Senthil/MSDataAnalytics/Semester1/Assignments/IS607/Week9Assignment/books1.html"
books1.html.data <- readHTMLTable(books1.html.url,header = T)
books1.html.data.df <- as.data.frame(books1.html.data)
str(books1.html.data.df)
print(books1.html.data.df)

```


### books.xml 
```{r}
#books.html.url <- getURL("https://raw.githubusercontent.com/senthiltamil/CUNY_MSDA/master/Week9_Assignment/books.xml")
books.xml.url <- "C:/Senthil/MSDataAnalytics/Semester1/Assignments/IS607/Week9Assignment/books.xml"
books.xml.data <- xmlParse(books.xml.url)
books.xml.list <- xmlToList(books.xml.data)
books.xml.df <- ldply(books.xml.list, data.frame)
str(books.xml.df)
print(books.xml.df)

```


### books.json
```{r}
#books.html.url <- getURL("https://raw.githubusercontent.com/senthiltamil/CUNY_MSDA/master/Week9_Assignment/books.json")
books.json.url <- "C:/Senthil/MSDataAnalytics/Semester1/Assignments/IS607/Week9Assignment/books.json"
books.json.data <- fromJSON(txt=books.json.url)
books.json.df1 <- as.data.frame(books.json.data)
str(books.json.df1)
print(books.json.df1)


```

###Conclusion:- All these format resulted in slightly different format of data frames. Json format was the easiest one to translate to dataframe using R.