---
title: "Week6Assignment"
output: html_document
---

##Non-graphical representation of numerical data

Install necessary packages
```{r}
#install.packages("ggplot2")
require(ggplot2)
```


1.a) Load demographic data from file. 

```{r}
file.path <- '/temp/demographicdata.csv'
demographic.data <- read.table(file.path, header=TRUE,sep=",")
```

1.b) structure of demographic data  is 

```{r}
# School.Name and School.Year are factors
# Grade.1, Grade.2, Grade.3, Grade.4 and Grade.5 are numerical variables.
str(demographic.data)
```

1.c) Top 15 rows to show the sample of data

```{r}
head(demographic.data, n=15)
```

1.d) Last 15 rows to show the sample of data

```{r}
tail(demographic.data, n=15)
```

2. Statistical Summary of data

```{r}
summary(demographic.data)
```


3. Frequency by School Name

```{r}
table(demographic.data$School.Name)
```

4. Frequency by School Year over School Name

```{r}
table(demographic.data$School.Year, demographic.data$School.Name)
```

-> sum of enrollment by School Year

```{r}
aggregate((Grade.1 + Grade.2 + Grade.3 + Grade.4 + Grade.5) ~ School.Year, data=demographic.data, sum)
```

-> sum of enrollment by  School Name

```{r}
aggregate((Grade.1 + Grade.2 + Grade.3 + Grade.4 + Grade.5) ~ School.Name, data=demographic.data, sum)
```

##Graphical representation of numerical data

5) Create different kinds of graphs for Grade1 data (single numeric value)

5.a) boxplot function from base r package

```{r}
boxplot(demographic.data$Grade.1)
```

5.a) histogram function from base r package

```{r}
hist(demographic.data$Grade.1)
```

5.c) qplot  function from ggplot2 package 

```{r}
qplot(demographic.data$Grade.1)
```

6) Create scatterplot graph for Grade1 and Grade2 data (Two numeric values)

6.a) Plot function from base r package

```{r}
plot(demographic.data$Grade.1 ~ demographic.data$Grade.2)
```

6.b) qplot function from ggplot2 package
```{r}
qplot(Grade.1, Grade.2, data=demographic.data)
```
