---
title: "Week12Assignment"
author: "Senthil Dhanapal"
date: "Thursday, April 23, 2015"
output: html_document
---

### I use rmongodb to read and display some information from local host mongo db. Data is inserted directly using mongo shell. There is a text file MongoDataManipulation.txt that shows the data manipulation commands and the test data

## Install necessary packages and load them
```{r, warning=FALSE}

#install.packages("rmongodb")
library(rmongodb)

```

## Connect to local host and check the connection status

```{r}

m<-mongo.create()
mongo.is.connected(m)


```

## list the collection of databases 

```{r}

mongo.get.databases(m)

```

## Get a list of collection from the test db, where the data is stored.

```{r}

mongo.get.database.collections(m, db = "test")

```


## Get the count of documents in my collection
```{r}

ns <- "test.inventory"
mongo.count(m, ns = ns)

```



## Get all the documents from the collections. 
## Display the count of documents fetched
## Display some information from the second document fetched
```{r}


find_all <- mongo.find.all(m, ns = ns)
sprintf("Records count in the collection is %i", NROW(find_all))
sprintf("For Item %s:- Manufacturer is %s and Category is %s", find_all[[2]]$item, find_all[[2]]$details$manufacturer, find_all[[2]]$category)


```

