---
title: "Project4"
author: "Senthil Dhanapal"
date: "Tuesday, April 21, 2015"
output: html_document
---
### I am using rvest library to scrape "http://www.r-bloggers.com/search/web%20scraping" to collect Date, author and title information. As additional task, I am navigating through all the pages and getting Date, author and title information from all those pages also.


### include necessary packages
```{r, warning=FALSE, message=FALSE}
#install.packages("rvest")
library(rvest)
library(stringr)
library(dplyr)

```



### A function rbloggerscraper is used to get the Date, title and author information. It takes url and pageid as parameter. This function is useful as it comes handy when I am navigating through all the pages in the pagination list.


```{r}

rbloggerscraper <- function(url, pageid)
{
  df.scrapedinfo <- data.frame(pageid=numeric(0), blogdate=character(),author=character(),title=character())
  rbloggerpage<-html(url)
  leftcontent <- rbloggerpage %>% html_nodes("#leftcontent") 
  posts <- leftcontent %>% html_nodes(xpath="//div[starts-with(@id,'post-')]")
  for(i in 1:length(posts))
  {
    blogdate <- html_nodes(posts[i],".meta .date") %>% html_text()
    author <- html_nodes(posts[i],".meta a") %>% html_text()
    title <- html_nodes(posts[i],"h2 a") %>% html_text()
    df.scrapedinfo <- rbind(df.scrapedinfo, data.frame(pageid,blogdate,author,title))
  }
  return (df.scrapedinfo)
}

```


### Get the data for the first page.
```{r}
df.result <- data.frame(pageid=numeric(0),blogdate=character(),author=character(),title=character())
df.r <- rbloggerscraper("http://www.r-bloggers.com/search/web%20scraping",1)
df.result <- rbind(df.result, df.r)

```


### Get the pagination information and get the lastpage number
```{r}

getLastPage<-function(url)
{
  pagination <- html(url) %>% html_nodes(".pagination")
  lastpage <- html_nodes(pagination,"a") %>% .[[5]] %>% html_text() %>% as.numeric()
  return(lastpage)
}

lastpage <- getLastPage("http://www.r-bloggers.com/search/web%20scraping")


```


### loop through pages from 2 to last page.
```{r}
scrapeRestOfPages <- function(partURL)
{
  df.r1 <- data.frame(pageid=numeric(0), blogdate=character(),author=character(),title=character())
  for(i in 2:lastpage)
  {
    url <-  str_c(partURL,i,"/")
    print(sprintf("Page number %i",i))
    df.r <- rbloggerscraper(url,i)
    df.r1 <- rbind(df.r1, df.r)
    Sys.sleep(2)
  }
  return(df.r1)
}


df1 <- scrapeRestOfPages("http://www.r-bloggers.com/search/web%20scraping/page/")
df.result <- rbind(df.result, df1)


```


###View the result data frame
```{r}
nrow(df.result)
tbl_df(df.result)

```


###Trying another additional task scraping twitter search

```{r}

df.result <- data.frame(pageid=numeric(0),blogdate=character(),author=character(),title=character())
df.r <- rbloggerscraper("http://www.r-bloggers.com/search/twitter",1)
df.result <- rbind(df.result, df.r)

lastpage <- getLastPage("http://www.r-bloggers.com/search/twitter")

df1 <- scrapeRestOfPages("http://www.r-bloggers.com/search/twitter/page/")
df.result <- rbind(df.result, df1)


```


```{r}

nrow(df.result)
tbl_df(df.result)

```

